---
format: 
  gfm:
    fig-format: png
knitr:
  opts_chunk:
    fig.path: "man/figures/README-"
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)

if (!requireNamespace("animovement", quietly = TRUE)) {
   devtools::load_all()
} else {
   library(animovement)
}
```


# animovement
<!-- badges: start -->

[![DOI](https://zenodo.org/badge/773406370.svg)](https://zenodo.org/doi/10.5281/zenodo.13235277)
[![R-CMD-check](https://github.com/animovement/animovement/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/animovement/animovement/actions/workflows/R-CMD-check.yaml)
[![animovement status badge](https://animovement.r-universe.dev/animovement/badges/version)](https://animovement.r-universe.dev/animovement)

<!-- badges: end -->

*An R toolbox for analysing movement across space and time*

The primary aim of the *animovement* package is to provide a unified, standardised workflow for analysing movement data in a *tidyverse*-friendly syntax. 

*We work actively with the developers of the Python [`movement`](https://movement.neuroinformatics.dev/) package, to reach a similar data standards, workflow and use cases; if you prefer analysing your data in Python, we highly recommend using [`movement`](https://movement.neuroinformatics.dev/).*

## Installation

You can install the development version of *animovement* with:

```{r, eval=FALSE}
install.packages('animovement', repos = c('https://animovement.r-universe.dev', 'https://cloud.r-project.org'))
```

Once you have installed the package, you can load it with:

```{r, eval=FALSE}
library("animovement")
```

## Documentation

Analysis of animal movement follows a similar workflow irrespective of the type of data (e.g. pose estimation, centroid tracking, trackball, treadmill). See our docs to go through the steps, one-by-one:

- [Introduction to `animovement`](https://animovement.dev/animovement/articles/animovement.html)
- [Read trackball data](https://animovement.dev/animovement/articles/read-trackball.html)
- [Clean tracks](https://animovement.dev/animovement/articles/clean-tracks.html)
- [Calculate kinematics](https://animovement.dev/animovement/articles/calculate-kinematics.html)
- [Calculate summary statistics](https://animovement.dev/animovement/articles/calculate-summary-statistics.html)

## Status
> **Warning**
>
> 🏗️ The package is currently in early development and the interface is subject to change. Feel free to play around and provide feedback.


## Contribute

**If your favourite type of movement data is not currently supported, we would love to get a sample of your data to support it!** 

If you enjoy the package, please make sure to [cite it](#citation). If you find a bug, feel free to open an issue!

<!-- ## Acknowledgements -->

<!-- *animovement* is all about the data, and we are deeply grateful for all those who have shared data with us to implement and test our code. Thank you! -->

<!-- - [Stan Edwards](): Trackball with optical flow, free. -->
<!-- - [Estelle Moubarak](): Trackball with optical flow, fixed. -->
<!-- - [Maria Cozan](): Treadmill with rotary encoder. -->
<!-- - [Violette Chiara](): AnimalTA -->

## Citation {#citation}

To cite *animovement* in publications use:

```{r}
citation("animovement")
```
