---
title: "Introduction to animovement"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to animovement}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# The startup banner is styled with ANSI escapes, which do not survive the
# trip into a rendered vignette.
options(animovement.styling = FALSE)
```

*animovement* is a meta-package. It does not analyse movement data itself — it
installs, attaches and keeps in sync the suite of packages that do, so that a
single `library(animovement)` gives you a complete and coherent workflow instead
of seven packages to track separately. In that respect it is a descendant of the
[*fastverse*](https://fastverse.github.io/fastverse/), whose approach to package
management it borrows.

Splitting the ecosystem into focused packages keeps each one small, testable and
useful on its own; bundling them into a meta-package means you rarely have to
think about that split. You get the whole workflow, at versions that are known
to work together, and a handful of functions for the housekeeping — which is
what this vignette is about. For tutorials on actually *analysing* movement data
with these packages, see [animovement.dev](https://animovement.dev).

## Attaching the animovement

```{r}
library(animovement)
```

Attaching prints a startup message listing the core packages and their versions,
followed by any conflicts between them and packages you already have loaded.
Both can be silenced with `options(animovement.quiet = TRUE)`.

The core packages are:

| Package | Role |
| --- | --- |
| [*aniframe*](https://animovement.dev/anicore) | Core data structures for movement data |
| [*aniread*](https://animovement.dev/aniread) | Reading and writing movement data |
| [*anispace*](https://animovement.dev/anispace) | Spatial transformation methods |
| [*aniprocess*](https://animovement.dev/aniprocess) | Signal processing and filtering |
| [*anicheck*](https://animovement.dev/anicheck) | Diagnosing movement data quality |
| [*animetric*](https://animovement.dev/animetric) | Calculating movement-based metrics |
| [*anivis*](https://animovement.dev/anivis) | Visualising movement data and diagnostics |

`animovement_packages()` reports the set that is actually in force — the core
packages, or whatever a project configuration file specifies, plus any session
extensions:

```{r}
animovement_packages()
```

## Conflicts

When two attached packages export a function of the same name, the one attached
later masks the other. `animovement_conflicts()` lists every such clash
involving an animovement package, in search-path order:

```{r}
animovement_conflicts()
```

It is also a general-purpose tool — pass it any set of attached packages to
check those instead:

```{r, eval=FALSE}
animovement_conflicts(sub("package:", "", search()[-1]))
```

## Extending the animovement

`animovement_extend()` adds packages for the current session. They are attached
alongside the core packages, their conflicts are reported, and they are
remembered in `options("animovement.extend")` so `animovement_packages()`,
`animovement_update()` and friends include them:

```{r, eval=FALSE}
animovement_extend(ggplot2, tidyr)
```

Use `install = TRUE` to install anything missing first. To extend *before*
attaching, set the option directly:

```{r, eval=FALSE}
options(animovement.extend = c("ggplot2", "tidyr"))
library(animovement)
```

`animovement_detach()` is the inverse. With no arguments it detaches all
animovement packages; `session = TRUE` also clears them from the session
options, and `unload = TRUE` unloads the namespaces as well:

```{r, eval=FALSE}
animovement_detach(ggplot2)                     # one package
animovement_detach(session = TRUE)              # everything, for good
```

## Project configuration

For a set of packages that belongs to a project rather than a session, put a
file named `.animovement` in the project root. List the packages one per line,
or separated by commas or spaces:

```
aniframe, aniread, aniprocess
ggplot2
```

When this file is present, it *replaces* the standard set of core packages — so
list everything the project needs. `animovement_packages()` reads it, and
`library(animovement)` attaches exactly that set.

Options and environment variables can be set from the same file by prefixing
them with `_opt_` (options) or giving them bare (environment variables). Entries
placed before the package names are applied before the packages are attached,
those after them afterwards:

```
_opt_animovement.install = TRUE
aniframe, aniread, aniprocess
```

The options the meta-package recognises are:

- `animovement.quiet` — `TRUE` disables startup, conflict and installation
  messages
- `animovement.styling` — `FALSE` disables styling of console output
- `animovement.extend` — packages to attach in addition to the core set
- `animovement.install` — `TRUE` installs missing packages on attach

## Installing and updating

The animovement packages are published on
[R-universe](https://animovement.r-universe.dev) rather than CRAN, so a plain
`getOption("repos")` will not find them. `animovement_repos()` adds the
R-universe to a repository vector, and is the default for the functions below:

```{r}
animovement_repos()
```

```{r, eval=FALSE}
options(repos = animovement_repos()) # ...or apply it to the whole session
```

`animovement_update()` checks the installed versions against the repositories
and installs what is behind; `animovement_install()` installs packages that are
missing altogether:

```{r, eval=FALSE}
animovement_update()
animovement_install()
```

`animovement_sitrep()` gives the situation report the two are based on — every
package in your animovement, its installed version, and whether a newer one is
available — and `animovement_deps()` lists the dependencies underneath them:

```{r, eval=FALSE}
animovement_sitrep()
animovement_deps()
```

## Suggested packages

To keep installation light, functionality that only some users need is a soft
dependency of the ecosystem packages rather than a hard one. Rather than
installing them one error message at a time, you can take them all at once:

```{r, eval=FALSE}
animovement_show_suggested()    # list them
animovement_install_suggested() # install them
```

Packages needed only for development or documentation (knitr, testthat, and the
like) are excluded, as are packages an ecosystem package already requires.
