---
title: "Calculate kinematics"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(animovement)
library(tibble)
library(ggplot2)
library(tidyr)
library(dplyr, warn.conflicts = FALSE)
library(here)
here::i_am("vignettes/articles/calculate-kinematics.Rmd")
```

```{r include=FALSE}
filepaths <- get_sample_data("trackball", quiet = TRUE)

df <- read_trackball(
  filepaths,
  setup = "of_free",
  col_time = "datetime",
  sampling_rate = 60
) |>
  # Trackball coordinates come out in sensor dots; calibrate to cm
  # (394 dots per cm for this mouse sensor).
  set_unit_space(to_unit = "cm", calibration_factor = 1 / 394)

# Smooth tracks
df_smooth <- df |>
  filter_aniframe(method = "rollmedian", window_width = 5, use_derivatives = TRUE)
```

## Calculate kinematics

When we work with movement data, we are often interested in more than just where an animal *is*; we're interested in how fast it moves, where it is heading etc. 
`calculate_kinematics` calculates a range of kinematic variables:

- `distance`: The distance the animal moved since the last observation (simply calculated using Pythagoras' theorem)
- `v_translation`: The translational velocity, like what you see on a speedometer in a car.
- `direction`: The direction (in radians) the animal is heading - where the arrow on the compass is heading.
- `rotation`: Difference from direction of the last observation.
- `v_rotation`: The rotational velocity (in rad/s).

```{r}
# Augment all data in list
df_kinematics <- df_smooth |>
  calculate_kinematics()
glimpse(df_kinematics)
```

### Clean kinematics

```{r}
df_kinematics_clean <- df_kinematics |>
  filter_na_speed(threshold = 100)
```

<!-- TODO: Commented out for now — these chunks use kinematics columns from the
previous calculate_kinematics() output (v_translation, v_rotation, direction,
d_rotation). The output schema has changed (speed, angular_velocity, heading,
angular_speed, …); re-enable once these plots are remapped to the new columns.

Alternatively, you can perform your own data cleaning methods.

```r
df_kinematics_alternative <- df_kinematics |>
  filter(v_translation > 0.25 & v_rotation < 200)
```

Whatever way you choose is all up to you, but do make sure you do not just calculate summaries without knowing what's in it.

Kinematics are much more prone to small values change. We assess these in two ways:

```r
library(patchwork)
a <- df_kinematics_clean |>
  ggplot(aes(x, y, colour = time)) +
  geom_path() +
  scale_colour_viridis_c()
b <- df_kinematics_clean |>
  ggplot(aes(time, direction)) +
  geom_line()
c <- df_kinematics_clean |>
  ggplot(aes(time, v_translation)) +
  geom_line()
d <- df_kinematics_clean |>
  ggplot(aes(time, v_rotation)) +
  geom_line()
e <- df_kinematics_clean |>
  ggplot(aes(abs(v_translation))) +
  geom_histogram()
f <- df_kinematics_clean |>
  ggplot(aes(abs(v_rotation))) +
  geom_histogram()
(a + b) / (c + d) / (e + f)

df_kinematics_clean |>
  tidyr::drop_na(d_rotation) |>
  select(d_rotation)
```

```r
pl_direction <- df_kinematics_clean |>
  ggplot(aes(x = direction, y = v_translation)) +
  stat_density_2d(
    geom = "tile",
    aes(fill = after_stat(density)),
    n = c(40, 10),
    contour = F
  ) +
  scale_fill_viridis_c() +
  coord_polar(
    direction = -1,
    start = -pi / 2
  ) +
  theme_minimal()

max_coord <- max(abs(c(df_kinematics_clean$x, df_kinematics_clean$y)), na.rm = TRUE)
pl_path <- df_kinematics_clean |>
  ggplot(aes(x, y, colour = v_translation)) +
  geom_path() +
  coord_fixed() +
  scale_x_continuous(limits = c(-max_coord, max_coord)) +
  scale_y_continuous(limits = c(-max_coord, max_coord)) +
  scale_colour_viridis_c() +
  theme_minimal()

pl_path + pl_direction +
  plot_layout(guides = "collect")
```
-->

